Characterizing subarctic biomes for land surface modeling of pollution and climate risk

### 2021-05-10 VegOzone.tex
Change naming from Norway spruce to coniferous

The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, downy birch, and Norway spruce. Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following. A comprehensive list of model parameters can be found in Supplement~Table~S1.
Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended warm period in late July did we find substantial stomatal conductance. In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during a short, subarctic growing season. Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation. Resulting response functions for Norway spruce (Picea abies) and downy birch are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

@@

For an acclimation with respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for downy birch and Norway spruce refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

\begin{table}[t]
  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Downy birch} & MM & 5 & 20 & 100\\
    & Boreal & 0 & 10 & 100\\
    & Cold & 5 & 15 & 100\\
    \middlehline
    \multirow{3}{*}{Norway spruce} & MM & 0 & 20 & 100\\
    & Boreal & 0 & 10 & 100\\
    & Cold & 0 & 15 & 100\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
    & Boreal & 0 & 10 & 24 \\
    & Cold & 5 & 15 & 36\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llr}
    \tophline
    Species & type & $\alpha$\\
    \middlehline
    \multirow{3}{*}{Downy birch} & MM & 0.004 \\
    & PPFD0.8 & 0.005\\
    & PPFD1.2 & 0.004\\
    \middlehline
    \multirow{3}{*}{Norway spruce} & MM & 0.006 \\
    & PPFD0.8 & 0.008  \\
    & PPFD1.2 & 0.005  \\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 0.011\\
    & PPFD0.8 & 0.014 \\
    & PPFD1.2 & 0.009 \\
    \bottomhline
    \end{tabular}
\end{table}

As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To quantify the resulting bespoke parameterizations, we propose the following metric. For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time). We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency. Our arbitrary optimization target is $60\,\unit{\%}$. Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions. For simplicity, we neglect the dependency on SWP in this regard. From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike. Overall, Norway spruce displays the least deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial.

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) Downy birch; (b) Norway spruce; (c) perennial grassland. For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Another important factor for ozone exposure metrics is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and is the value used for all PFTs. The resulting growing season for Norway spruce in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
For comparison, we would get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ from the latitude model typically used for localizing $\mathrm{DO_3SE}$. By comparison, the latitude model indicates about a $2\,\unit{month}$ longer growing season. Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.

\begin{table}[t]
  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_gs}
  \begin{tabular}{llcc}
    \tophline
    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
    \middlehline
    \multirow{2}{*}{Downy birch} & 2018 & 129$^*$ & 261$^\dagger$ \\
    & 2019 & 130$^*$ & 274$^\dagger$ \\
    \multirow{2}{*}{Norway spruce} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
\end{table}

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, downy birch, Norway spruce, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown in Fig.~\ref{fig:pody_rel}. For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce) \citep{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines. Different years are color coded as previously. Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles. The annotation $\pm$ refers to bespoke light response function threshold ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

@@

For Norway spruce (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Norway spruce is the least sensitive to an acclimation in temperature and therefore the least biased compared to the generic default MM parameterization. Norway spruce displays a medium sensitivity to the variation in the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019). (a) Downy birch; (b) Norway spruce; (c) Perennial grassland. The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{boreal}) for each bespoke PFT. The estimates are listed in Table~\ref{tab:biomass_reduction}. The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years. Norway spruce is the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties. This may reflect the more comprehensive research on this species or PFT. The systematic deviation for downy birch is one order of magnitude larger than for Norway spruce and perennial grassland.

@@

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{boreal} with bespoke GS and SWP=off. The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on. For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown. The reported standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and bespoke GS).}
  \label{tab:biomass_reduction}
\begin{tabular}{lllll}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Downy birch & Norway spruce & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
\middlehline
$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

@@

We found that the length and timing of the growing season are crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$). Based on a MODIS photosynthesis product, we estimated the growing season for Norway spruce in 2018 to be at least one month shorter than in 2019. The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}. Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively. The timing and emerging trend in the start of the growing season may differ substantially between local observation and common thermal or latitudinal estimations. Species-specific methods show a later start for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}. With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd.

>>>>

The main species or PFTs, we deduced on-site at the NIBIO Environment Centre Svanhovd and from Fig.~\ref{fig:svanhovd_research_station}, are perennial grassland, birch, and pine. Default parameters for these species or PFTs are derived from \citet{EP:Simpson2007,GCB:Mills2011,ICP:MappingManual2017} and will be referred to as \emph{mapping manual} abbreviated as MM in the following. We base our parameterization of coniferous trees on boreal Norway spruce from the MM. Birch will be referred to as deciduous tree in the following. A comprehensive list of model parameters can be found in Supplement~Table~S1.
Initial test simulations with the MM PFTs showed unrealistically low $g_\mathrm{sto}$ in 2019 particularly for perennial grassland (Appendix~Fig.~\ref{fig:pody_mm_composit}). Only during an extended warm period in late July did we find substantial stomatal conductance. In ecological terms this would mean that there was almost no growth of grass in the summer of 2019 -- a prediction which is easily falsified by reality. We identified $f_\mathrm{T}$ as abnormally low, being the limiting factor of stomatal conductance. Therefore, we propose bespoke parameterizations which may describe an actual acclimation of these species to the climatic conditions in the target region.

@@

We hypothesize that perennial vegetation will likely prioritize their growth or fast adolescence and hence their carbon sequestration, during a short, subarctic growing season. Because the mapping manual version of the $\mathrm{DO_3SE}$ model currently does not simulate $A_\mathrm{n}$, we assume a first order proportionality between $A_\mathrm{n}$ and $g_\mathrm{sto}$ \citep{GCB:Medlyn2011} and effectively tune temperature and light response functions ($f_\mathrm{T}$ and $f_\mathrm{light}$) for higher $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. We further assume that $f_\mathrm{VPD}$ and $f_\mathrm{SWP}$ suffice our biomes. The procedure is graphically illustrated in Fig.~\ref{fig:f_temp_grassland} for grassland. We aim to increase the enclosed area between probability density function (PDF) and the respective response function. We show two climatological reference periods 1990s and 2000s to indicate that the vegetation is subject to ongoing climate change and hence continuous acclimation. Resulting response functions for coniferous and deciduous trees are displayed in Appendix~Figs.~\ref{fig:f_temp_spruce}--\ref{fig:f_temp_birch}.

@@

For an acclimation with respect to light conditions, we hypothesize that the light threshold for a $50\,\unit{\%}$ opening of plants' stomata may differ in subarctic species. We analytically derive the inverse function $f_{\mathrm{light},k}^{-1}$ for each PFT or species $k$
\begin{equation}
  \gamma_k(f_\mathrm{light}) := f_{\mathrm{light}, k}^{-1} = -\frac{\ln(1-f_\mathrm{light})}{\alpha_k}.
  \label{eq:inverse_function}
\end{equation}
First, we solve Eq.~(\ref{eq:inverse_function}) for the MM default value $\alpha_\mathrm{MM}^k$ for each PFT or species $k$ at $0.5$ 
\begin{equation}
  \gamma^k(0.5) = -\frac{\ln(0.5)}{\alpha_\mathrm{MM}^k},
  \label{eq:inverse_function_halfway}
\end{equation}
and define a variation $\gamma' = \gamma \cdot \eta$ with $\eta \in \{0.8, 1.2\}$. We invert Eq.~(\ref{eq:inverse_function}) for $\alpha$ and find $\alpha'(\gamma')$. The resulting functions for perennial grassland are shown in Fig.~\ref{fig:f_temp_grassland}b) (for deciduous and coniferous trees refer Appendix~Figs.~\ref{fig:f_temp_spruce}b)--\ref{fig:f_temp_birch}b)). All derived parameters are tabulated in Table~\ref{tab:sensitivity_tests_light}.

@@

\begin{table}[t]
  \caption{Bespoke temperature parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_temp}
  \begin{tabular}{llrrr}
    \tophline
    Species & type & $T_\mathrm{min}$ & $T_\mathrm{opt}$ & $T_\mathrm{max}$ \\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 5 & 20 & 100\\
    & Boreal & 0 & 10 & 100\\
    & Cold & 5 & 15 & 100\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0 & 20 & 100\\
    & Boreal & 0 & 10 & 100\\
    & Cold & 0 & 15 & 100\\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 10 & 24 & 36\\
    & Boreal & 0 & 10 & 24 \\
    & Cold & 5 & 15 & 36\\
    \bottomhline
    \end{tabular}
\end{table}

\begin{table}[t]
  \caption{Bespoke light parameterizations. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_light}
  \begin{tabular}{llr}
    \tophline
    Species & type & $\alpha$\\
    \middlehline
    \multirow{3}{*}{Deciduous tree} & MM & 0.004 \\
    & PPFD0.8 & 0.005\\
    & PPFD1.2 & 0.004\\
    \middlehline
    \multirow{3}{*}{Coniferous tree} & MM & 0.006 \\
    & PPFD0.8 & 0.008  \\
    & PPFD1.2 & 0.005  \\
    \middlehline
    \multirow{3}{*}{Perennial grassland} & MM & 0.011\\
    & PPFD0.8 & 0.014 \\
    & PPFD1.2 & 0.009 \\
    \bottomhline
    \end{tabular}
\end{table}

As indicated above, a better acclimation will result in a higher average $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$. To quantify the resulting bespoke parameterizations, we propose the following metric. For a generic GS (May--August), we compute climatological average and standard deviation of $\left<g_\mathrm{sto}\right>/g_\mathrm{max}$ (Eq.~(\ref{eq:stomatal})) at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time). We presume that relative stomatal conductance of around noon (highest light intensity) and in the early morning (lower intensities) are good proxies for \chem{CO_2} uptake efficiency. Our arbitrary optimization target is $60\,\unit{\%}$. Additional constrains are a small deviation between noon and morning and a low standard deviation indicating a higher robustness to interannual variability of growing conditions. For simplicity, we neglect the dependency on SWP in this regard. From the results shown in Fig.~\ref{fig:javis_func_opt_mean}, we identify boreal, PPFD0.8 as the best acclimation for all hypothetical species or PFTs alike. As expected due to the small adjustments, coniferous trees display the lowest deviation between the different types of parameterizations, while the divergence for perennial grassland is substantial as response to the stronger proposed temperature acclimation.

\begin{figure}[t]
  %\includegraphics[width=8.3cm]{javis_func_opt_mean}
  \includegraphics[width=8.3cm]{fig10}
  \caption{Proposed metric to test bespoke response functions. GS (May--August) climatological averages and standard deviation of $g_\mathrm{sto}^k$ (Eq.~(\ref{eq:stomatal})) relative to $g_\mathrm{max}^k$ at noon (averaged over $11\,\unit{am}-1\,\unit{pm}$ local time) and in the morning (averaged over $5-9\,\unit{am}$ local time) are shown for (a) deciduous tree; (b) coniferous tree; (c) perennial grassland. For simplicity, $f_\mathrm{SWP}$ is neglected in this calculation. }
  \label{fig:javis_func_opt_mean}
\end{figure}

Another important factor for ozone exposure metrics is the timing and length of the GS. Start ($A_\text{start}$) and end ($A_\text{end}$) of GS for each PFT are given in units of \unit{doy} (Tab.~\ref{tab:sensitivity_tests_gs}) and estimated as follows. Evergreen needle leaf trees show photosynthetic activity already at low temperatures \citep{TP:Wallin2013, TB:Kolari2007}. We use a MODIS (Aqua/Terra) retrieved net photosynthesis product for a $1\times 1\,\unit{km}$ patch centered at Svanhovd to determine $A_\text{start}$ and $A_\text{end}$. Net photosynthesis data can be approximated by a second order polynomial function (Fig.~\ref{fig:modis_Psn}). By calculating the roots of the fitted polynomial, we derive $A_\text{start} = 122\,\unit{doy}$ for 2018 and \unit{doy} 106 in 2019. $A_\text{end}$ amounts to \unit{doy} 261 and 274, respectively, and is the value used for all PFTs. The resulting growing season for coniferous trees in 2019 was one month longer then in 2018. For estimating $A_\text{start}$ of downy birch, we use the $5$ consecutive days above $5\,\unit{^\circ C}$ agricultural rule of thumb on gridded temperature data from SeNorge.no (Fig.~\ref{fig:greening_season_change_Svanvik}). We find 129 and 130, respectively. For both years, these dates coincident with the first snow-free day at the meteorological weather station at Øvre Neiden (Sør-Varanger, NOR) which is relatively close-by.
For comparison, we would get $A_\mathrm{start} = 100\,\unit{doy}$ and $A_\mathrm{end} = 307\,\unit{doy}$ from the latitude model typically used for localizing $\mathrm{DO_3SE}$. By comparison, the latitude model indicates about a $2\,\unit{month}$ longer growing season. Due to a lack of quantitative field observation, we assume a latency period of about $1\,\unit{month}$ after snow-melt for perennial grassland.

\begin{table}[t]
  \caption{Bespoke start and end of growing season in units of \unit{doy}. MM refers to mapping manual \citep{GCB:Mills2011,ICP:MappingManual2017}.}
  \label{tab:sensitivity_tests_gs}
  \begin{tabular}{llcc}
    \tophline
    Species & Year & $A_\mathrm{start}$ & $A_\mathrm{end}$\\
    \middlehline
    \multirow{2}{*}{Deciduous tree} & 2018 & 129$^*$ & 261$^\dagger$ \\
    & 2019 & 130$^*$ & 274$^\dagger$ \\
    \multirow{2}{*}{Coniferous tree} & 2018 & 122$^\dagger$ & 261$^\dagger$ \\
    & 2019 & 106$^\dagger$ & 274$^\dagger$ \\
    \multirow{2}{*}{Perennial grassland} & 2018 & 159$^\ddagger$ & 261$^\dagger$\\
    & 2019 & 161$^\ddagger$ & 274$^\dagger$ \\
    \bottomhline
  \end{tabular}
  \belowtable{$^\dagger$ MODIS (Aqua/Terra) net photosynthesis product;\\
    $^*$ $5\,\unit{days}-5\,\unit{^\circ C}$-rule, $T_\mathrm{air}$ from \href{seNorge.no}{seNorge.no};\\
    $^\ddagger$ One month after snow melt; reference station Øvre Neiden (Sør-Varanger, NOR)} % Table Footnotes
\end{table}

@@

We have modeled $\mathrm{POD_y}$ with a flux threshold $y=1\,\unit{nmol\,m^{-2}\,s^{-1}}$ per projected leaf area (PLA) for three natural/semi-natural vegetation types, deciduous trees, coniferous trees, and perennial grassland using both MM parameterizations and a variety of bespoke response function in $f_\mathrm{T}$ and $f_\mathrm{light}$, and localized GS (described in detail in Section~\ref{subsec:do3se_parameters}). The results are shown in Fig.~\ref{fig:pody_rel}. For comparison, critical levels for a $4\,\unit{\%}$ (silver birch), $2\,\unit{\%}$ (Norway spruce) \citep{ICP:MappingManual2017}, and $10\,\unit{\%}$ (perennial grassland, above ground) \citep{ESPR:Hayes2021} reduction in biomass are shown as horizontal dashed lines. Different years are color coded as previously. Simulations with MM GS, canopy, and leaf parameters are indicated with squared markers, bespoke GS with circles. The annotation $\pm$ refers to bespoke light response function threshold ($+\rightarrow\mathrm{PPFD1.2}$, $-\rightarrow\mathrm{PPFD0.8}$), while bespoke temperature response functions are indicated by categories on the x-axis. Closed/open symbols represent SWP sensitivity switched off/on in the model.

@@

For coniferous trees (Fig.~\ref{fig:pody_rel}b), most simulations slightly exceed the CL ($0-10\,\unit{mmol\,\chem{O_3}\,m^{-2}}$). Coniferous trees' acclimation to prevailing temperatures was already good from the beginning and therefore the least biased of the generic MM parameterization. Coniferous trees display a medium sensitivity to the variation in the light threshold, e.g. $\Delta\mathrm{POD_1} = ^{+1.7}_{-1.3}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in case of boreal type without bespoke GS in 2018. Taking the bespoke GS into account, we find a higher ozone uptake in 2019 for all bespoke temperature response functions. The overall bias deduced from bespoke GS for boreal and cold types with respect to MM ranges between $(0.5-1.0)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2018 and $(2.0-3.3)\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ in 2019. The maximum systematic deviations deduced from the $-20\,\unit{\%}$ (superscript) and $+20\,\unit{\%}$ (subscript) variation of the extent of stomatal opening at low light intensities amount to $^{+1.0}_{-0.8}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$ and $^{+1.3}_{-1.0}\,\unit{mmol\,\chem{O_3}\,m^{-2}}$, respectively.

@@

\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig11}
  \caption{$\mathrm{DO_3SE}$ modeling results for $\mathrm{POD_1}$ with default parameterization (MM) and bespoke parameters for Svanhovd. Bespoke parameterizations include temperature response functions for boreal, cold and light thresholds (PPFD0.8, PPFD1.2) at the end of a MM GS (lighter hue and dotted lines) and the estimated GS for Svanhovd (2018, 2019). (a) Deciduous tree; (b) coniferous tree; (c) perennial grassland. The dashed horizontal lines indicate critical levels defined in \citet{ICP:MappingManual2017,ESPR:Hayes2021} for each species, respectively.}
  \label{fig:pody_rel}
\end{figure}

We estimated the biomass reductions in accordance to the CLs in \citet{ICP:MappingManual2017, ESPR:Hayes2021} for the most extreme temperature acclimation (\emph{boreal}) for each bespoke PFT. The estimates are listed in Table~\ref{tab:biomass_reduction}. The reported range in uncertainty corresponds to the maximum deviation defined by the $\pm 20\,\unit{\%}$ variation of the extent of stomatal opening at low light intensities with SWP=on. For downy birch and perennial grassland, substantial reductions in total biomass in comparison with the default generic MM parameterizations are predicted for both years. Coniferous trees are the least affected and the magnitude of total biomass reduction is independent of the choice of the parameterization within given uncertainties. This may reflect the more comprehensive research on this species or PFT feeding into the generic MM parameterization. The systematic deviation for deciduous trees is one order of magnitude larger than for coniferous trees and perennial grassland.

@@

\begin{table}[t]
  \caption{Estimated total biomass reduction in \unit{\%} for temperature acclimation \emph{boreal} with bespoke GS and SWP=off. The uncertainty ranges reported here correspond to maximum a divergence deducted from varying the extent of stomatal opening at low light intensities and SWP=on. For comparison, the corresponding biomass reduction for the generic default MM averaged over both years is shown. The reported standard deviation is computed from all sensitivity simulations ($\pm 20\,\unit{\%}$ in extent of stomatal opening at low light, SWP on/off, and bespoke GS).}
  \label{tab:biomass_reduction}
\begin{tabular}{lllll}
\tophline
Year & \multicolumn{4}{c}{PFT}\\
& Deciduous tree& Coniferous tree & \multicolumn{2}{c}{Perennial grassland}\\
\middlehline
2018 & $15.5^{+(1.9...2.1)}_{-(0.8...1.5)}$ & $2.5^{+0.2}_{-0.2}$ & $9.7^{+0.2}_{-0.2}$ & $^*12.4^{+0.2}_{-0.2}$\\
2019 & $17.4^{+2.5}_{-1.8}$ & $3.0^{+0.2}_{-0.3}$ & $10.5^{+(0.1...0.2)}_{-(0.0...0.1)}$ & $^*13.7^{+(0.1...0.3)}_{-(0.1...0.3)}$\\
\middlehline
$\left<\mathrm{MM}\right>$ & $11.2\pm 1.1$ & $2.31\pm 0.04$ & $7.5\pm 0.9$ & $^*9.6\pm 1.4$\\
\bottomhline
\end{tabular}
\belowtable{$^*$ above ground biomass} % Table Footnotes
\end{table}

@@

We found that the length and timing of the growing season are crucial for risk assessment based on accumulated ozone flux metrics ($\mathrm{POD_y}$). Based on a MODIS photosynthesis product, we estimated the growing season for coniferous trees in 2018 to be at least one month shorter than in 2019. The found start of growing season in 2018 ($122\,\unit{doy}$) and 2019 ($106\,\unit{doy}$) are within given ranges by empirical observations \citep{TB:Kolari2007,IVL:Karlsson2018}. Based on gridded observational temperature and snow depth, we found a half a month shorter growing season for downy birch and perennial grassland in 2018 compared to 2019, respectively. The timing and emerging trend in the start of the growing season may differ substantially between local observation and common thermal or latitudinal estimations. Species-specific methods show a later start for European silver birch and an earlier for Norway spruce, compared to the thermal growing season \citep{IVL:Karlsson2018}. With respect to ongoing climate change, a clear positive trend in length ($5.2\,\unit{d\,decade^{-1}}$) almost equally distributed between earlier start ($2.9\,\unit{days\,decade^{-1}}$) and later end ($2.3\,\unit{d\,decade^{-1}}$) could be deducted from the thermal growing season at Svanhovd.
